### 2019-11-01 design.tex
s1-3 and letter comments

\stitle{Why a Multi-granularity Binned Overview?}
\saj{A standard overview approach
is to show a spatially partitioned
collection of thumbnails \saj{on the left of the detailed view},
similar to Microsoft PowerPoint or
Adobe Reader.
However, displaying too many thumbnails
results in increased scrolling
to access distant thumbnails,
increasing visual discontinuity.
On the other hand,
displaying too few thumbnails
reduces visual discontinuity,
but at the cost of visual
clarity---the thumbnails appear cluttered
and fail to represent the
underlying data clearly~\cite{cockburn2009review}.
To strike a balance between these two objectives
({\bf DC3}) we designed a multi-granularity overview
that abstracts the data at varying levels of detail.
Multi-granularity interfaces
provide an alternative to \newsaj{the spatially partitioned single-granularity representation of the data space, \eg in Powerpoint, by allowing users to control the scale at which the overview should be displayed~\cite{cockburn2009review}.}
Users can resize the overview to control the amount of spreadsheet data that remains visible.
\toappendix{Microsoft PowerPoint also uses a
similar technique to balance the screen
space between overview and detail.}
Users can also hide the overview if required.}

\newsaj{The data structure underlying the overview
is an equi-depth histogram constructed
on the values
in the navigation attribute column.
Equi-depth histograms are commonly
used for summarizing the statistical properties
of data, with applications such as
database query optimizers~\cite{chaudhuri1998random}.
The binned representation of the overview
using equi-depth histograms
enables faster navigation
by reducing visual discontinuity. 
The bins act as landmarks
in the overview, enabling 
users to skip irrelevant bins
and quickly navigate to the desired subset of data. We now discuss how the overview is constructed.}

>>>>

\stitle{Why a Multi-granularity Binned Overview?}
\saj{A standard overview approach
is to show a spatially partitioned
collection of thumbnails \saj{on the left of the detailed view},
similar to Microsoft PowerPoint or
Adobe Reader.
However, displaying too many thumbnails
results in increased scrolling
to access distant thumbnails,
increasing visual discontinuity.
On the other hand,
displaying too few thumbnails
reduces visual discontinuity,
but at the cost of visual
clarity---the thumbnails appear cluttered
and fail to represent the
underlying data clearly~\cite{cockburn2009review}.
To strike a balance between these two objectives
({\bf DC3}) we designed a multi-granularity overview
that abstracts the data at varying levels of detail.
\newsaj{Multi-granularity representations 
have shown to
scale better to 
larger datasets----presenting information 
at multiple granularities
makes visual representations
more perceptually scalable
and less cluttered~\cite{elmqvist2009hierarchical}.
The multi-granularity overview of \noah
provides an alternative to the spatially partitioned single-granularity representation of the data space, \eg in Powerpoint, by allowing users to control the scale at which the overview should be displayed~\cite{cockburn2009review}.}
Users can resize the overview to control the amount of spreadsheet data that remains visible.
\toappendix{Microsoft PowerPoint also uses a
similar technique to balance the screen
space between overview and detail.}
Users can also hide the overview if required.}

\newsaj{The data structure underlying the overview
is an equi-depth histogram constructed
on the values
in the navigation attribute column.
Equi-depth histograms are commonly
used for summarizing the statistical properties
of data, with applications such as
database query optimizers~\cite{chaudhuri1998random}.
The binned representation 
of the overview
using equi-depth histograms 
aggregates data 
and visualizes density 
by counting 
the number of data points 
falling within each bin. 
Binned aggregation 
highlights both high level (\eg densities) and low level (e.g., outliers)
details of large datasets, 
while enabling
multi-granularity representation of data
via the choice of bin size~\cite{liu2013immens}. 
Moreover, binning reduces 
visual discontinuity during navigation
as users are able to 
view an overview 
that fits in the computer screen.
As a result, users can 
quickly navigate the data---the bins act as landmarks
in the overview, enabling 
users to skip irrelevant bins
and quickly navigate to the desired subset of data. We now discuss how the overview is constructed.}

### 2019-11-01 letter.tex
changes until related
s1-3 and letter comments

clarify the scope of tasks supported by 
a \noah-enhanced spreadsheet; our new Table~\ref{tab:scope}

@@

\item \textit{Explanation of the binning mechanism} (\textbf{C2}): 
\new{We now explain our choice
of the binning mechanism in 
the context of the third design consideration (\emph{DC3}), 
as detailed in Section~\ref{sec:design}.
We further expand on this in the blurb titled 
``Why a Multi-granularity Binned Overview'' 
in Section~\ref{sec:ui}.
In brief, we opted for binning to 
provide a clear and concise 
representation of the overall data 
distribution while minimizing user's back 
and forth movement across multiple screens 
during navigation. 
At some level, we do need to limit the number
of values displayed so that the overview
fits on the screen: binning is a natural solution
for that issue. 
\agp{Possibly argue using the immens paper?}
However, we acknowledge limitations 
for this choice for binning categorical data 
and discuss possible solutions 
in Section~\ref{sec:discuss} 
in the context of our user study findings. 
}

\item \textit{Justification of Excel as a Baseline} (\textbf{C2}):
\new{Since our goal is to improve spreadsheet
user experience---and there are very good reasons for doing so,
as outlined in item 1) above---the natural comparison point
is Excel. 
That said, beyond popularity, 
we have provided a thorough justification
for what spreadsheet systems like Excel offer
relative to tabular data analysis tools (TDA),
making it a more appropriate point of comparison.
Our justification for using Excel over 
tabular data analysis (TDA) tools involves 
a) highlighting the appeal of spreadsheets 
among users who shun more advanced enterprise 
tools (Section~\ref{sec:intro}), 
b) identifying the scope of tasks 
supported by spreadsheets and 
TDA tools (Section~\ref{sec:usage} and~\ref{sec:related}), 
c) explaining that the goal of spreadsheets 
is to present the raw data as is, 
amenable to editing, formula computation, and comparison of derived data, all done in-situ,
unlike TDA tools that hide the data 
while providing summarized statistics 
(Section~\ref{sec:related}), and 
d) highlighting the importance of 
addressing the shortcomings of a tool, 
the user base of which far exceeds 
that of TDA tools (Section~\ref{sec:study}). 
}

\item \textit{Enhancing the user study design} 
(\textbf{C4, C5}):
\agp{Revise to clarify where these changes are made
and HOW these changes solve the issue. Just saying
we merged doesn't necessarily address the issue.}
We have added additional 
explanations regarding the goal 
of the study while better articulating 
the research questions, provided a better justification for the choice of the tasks for the study (see Section~\ref{sec:study_design}) and added a detailed discussion on the limitations of the study. \saj{Specifically, we merged a number of research questions that explored how specific features of \noah affected participants' navigation experience into one broad question which explores their overall experience with \noah and its components. Moreover, we have justified the choice of the quiz tasks by a) relating their goals with the scope of tasks supported by spreadsheets (see Section~\ref{sec:usage}), and b) explaining how these tasks can reveal how people may use \noah for spreadsheet navigation. Finally, we have identified two limitations of the study, resulting from the sample size and the study design, which we justify based on the scope of the study.}

\item \textit{Reporting Intra-participant Differences} (\textbf{C6}):
\new{We performed additional analysis of the study data to report intra-participant submission time differences while providing insights on the trends that emerged from analyzing the data: this new analysis can be found in XXX\agp{pls add; insights on trends sounds vague let's be more clear}.} \saj{For the latter, we find that, despite a few exceptions, across all tasks, participants' task submission times were faster using \noah. However, we already obtained similar observations from Figure~\ref{fig:timeBox} in Section~\ref{sec:rq1}. Therefore, we have included the results of the intra-participant analysis in appendix while briefly summarizing its outcomes in Section~\ref{sec:rq1}.}
\end{enumerate}

>>>>

clarify the scope of tasks supported by spreadsheets and enhanced
by a spreadsheet with a \noah plug-in; our new Table~\ref{tab:scope}

@@

\item \textit{Explanation of the binning mechanism} (\textbf{C2}): 
\new{We now explain our choice
of the binning mechanism in 
the context of the third design consideration (\emph{DC3}), 
as detailed in Section~\ref{sec:design}.
We further expand on this in the blurb titled 
``Why a Multi-granularity Binned Overview'' 
in Section~\ref{sec:ui}.
In brief, we opted for binning to 
provide a clear and concise 
representation of the overall data 
distribution while minimizing user's back 
and forth movement across multiple screens 
during navigation. 
At some level, we do need to limit the number
of values displayed so that the overview
fits on the screen: binning is a natural solution
for that issue. 
\agp{Possibly argue using the immens paper?}
\newsaj{We cite related work on 
multi-scale aggregation~\cite{elmqvist2009hierarchical} 
to motivate the choice of 
a multi-granularity overview
and show that binned aggregation via histogram~\cite{liu2013immens}
is a suitable representation of such an overview.}
However, we acknowledge limitations 
for this choice for binning categorical data 
and discuss possible solutions 
in Section~\ref{sec:discuss} 
in the context of our user study findings. 
}

\item \textit{Justification of Excel as a Baseline} (\textbf{C3}):
\new{Since our goal is to improve spreadsheet
user experience---and there are very good reasons for doing so,
as outlined in item 1) above---the natural comparison point
is Excel. 
That said, beyond popularity, 
we have provided a thorough justification
for what spreadsheet systems like Excel offer
relative to tabular data analysis tools (TDA),
making it a more appropriate point of comparison.
Our justification for using Excel over 
tabular data analysis (TDA) tools involves 
a) highlighting the appeal of spreadsheets 
among users who shun more advanced enterprise 
tools (Section~\ref{sec:intro}), 
b) identifying the scope of tasks 
supported by spreadsheets and 
TDA tools (Section~\ref{sec:usage} and~\ref{sec:related}), 
c) explaining that the goal of spreadsheets 
is to present the raw data as is, 
amenable to editing, formula computation, and comparison of derived data, all done in-situ,
unlike TDA tools that hide the data 
while providing summarized statistics 
(Section~\ref{sec:related}), and 
d) highlighting the importance of 
addressing the shortcomings of a tool, 
the user base of which far exceeds 
that of TDA tools (Section~\ref{sec:study}). 
}

\item \textit{Enhancing the user study design} 
(\textbf{C4, C5}):
\agp{Revise to clarify where these changes are made
and HOW these changes solve the issue. Just saying
we merged doesn't necessarily address the issue.}
We have added additional 
explanations regarding the goal 
of the study while better articulating 
the research questions, provided a better justification for the choice of the tasks for the study (see Section~\ref{sec:study_design}) and added a detailed discussion on the limitations of the study. \newsaj{Specifically, we merged our previous \emph{RQ1} (evaluating the quiz task performance) with \emph{RQ4} (summarizing user feedback on performance via survey) into a unified research question (new \emph{RQ1}) on participants' navigation performance. Moreover, we merged a previous \emph{RQ2} and \emph{RQ3} that explored how specific features of \noah affected participants' navigation experience into one broad question (new \emph{RQ2}) which explores their overall experience with \noah and its components}. Moreover, we have justified the choice of the quiz tasks by a) relating their goals with the scope of tasks supported by spreadsheets (see Section~\ref{sec:usage}), and b) explaining how these tasks can reveal how people may use \noah for spreadsheet navigation. Finally, we have identified three limitations of the study, resulting from the sample size and the study design, which we justify based on the scope of the study.

\item \textit{Reporting Intra-participant Differences} (\textbf{C6}):
\newsaj{We performed additional analysis of the study data to report intra-participant submission time differences for the quiz phase tasks. While we summarize the results of the analysis in Section~\ref{sec:time_acc}, the detailed result can be found in Appendix}\agp{pls add; insights on trends sounds vague let's be more clear}. \newsaj{We find that, despite a few exceptions, across all tasks, participants' task submission times were faster using \noah---all but one participants completed at least four tasks in less time using \noah compared to Excel.} However, we already obtained similar observations from Figure~\ref{fig:timeBox} in Section~\ref{sec:rq1}. Therefore, we have included the results of the intra-participant analysis in Appendix.
\end{enumerate}

### 2019-11-01 results.tex
s1-3 and letter comments

\subsubsection{Faster navigation without sacrificing accuracy}
In Figure~\ref{fig:timeBox}a and~\ref{fig:timeBox}b, 
we show the distribution of submission times of participants 
for the four task categories, for birdstrikes and Airbnb respectively. 
For most categories, 
\emph{participants' median submission times using \noah 
were less than the fastest submission times using Excel}. This suggests that
the new capabilities offered by \noah
made spreadsheet navigation faster for these tasks. We also analyzed the intra-participant submission time difference, and that supported the aforementioned observations. Overall, majority of the submission times using \noah were faster compared to Excel---19 out of the 20 participants completed at least four tasks in less time using \noah compared to Excel. The submission time differences were more prominent for the steer, identify, and \cmpB tasks. Moreover, for all tasks except customize, the difference in submission times was statistically significant. Both the intra-participant difference and the statistical significance test results are discussed in detail in the Appendix.

>>>>

\subsubsection{Faster navigation without sacrificing accuracy}
\label{sec:time_acc}
In Figure~\ref{fig:timeBox}a and~\ref{fig:timeBox}b, 
we show the distribution of submission times of participants 
for the four task categories, for birdstrikes and Airbnb respectively. 
For most categories, 
\emph{participants' median submission times using \noah 
were less than the fastest submission times using Excel}. This suggests that
the new capabilities offered by \noah
made spreadsheet navigation faster for these tasks. \saj{We also analyzed the intra-participant submission time difference, and that supported the aforementioned observations. Overall, majority of the submission times using \noah were faster compared to Excel---19 out of the 20 participants completed at least four tasks in less time using \noah compared to Excel. The submission time differences were more prominent for the steer, identify, and \cmpB tasks. Moreover, for all tasks except customize, the difference in submission times was statistically significant. Both the intra-participant difference and the statistical significance test results are discussed in detail in the Appendix.}

### 2019-11-01 study.tex
s1-3 and letter comments

In this section, we present the design of a user study
to evaluate whether 
\noah helps address spreadsheet navigational challenges.
\subsection{Study Design and Participants}
\label{sec:study_design}
\newsaj{The goal of our study is to evaluate whether \noah addresses the navigational challenges in traditional spreadsheets. Furthermore, we wanted understand how the presence of the features introduced by a plugin like \noah impacts a spreadsheet user's navigation experience as they explore and make sense of data. Therefore, we decided to compare a \noah-integrated spreadsheet with a typical, popular one, Excel, for various navigational tasks in the spreadsheet domain.} \saj{Similar domain specific-evaluations have been performed for measuring the performance of various  overview+detail interfaces, \eg database browsing~\cite{north2000snap} or 
tree navigation~\cite{beard1990navigational}. As explained in Section~\ref{sec:related}, the goals \newsaj{and user populations} of spreadsheets and TDA tools are quite different. Therefore, we did not consider TDA tools for the comparative study.}
Our study was designed to answer the following questions:
\agp{The RQs feel repetitive. What did Karrie think.}
\newsaj{She was fine with it. I think there are subtle differences. the RQ1 tries corresponds to quantifiable performance metric like accuracy, latency. RQ2 is qualitative feedback of user experience.}
\squishlist
\item {\bf RQ1.} %\sout{How does the addition of an overview help address spreadsheet navigational challenges, with respect to quantitative performance metrics such as speed and accuracy?}
\saj{How did the addition of an overview like \noah impact \newsaj{a user's spreadsheet navigation performance  measured in terms of speed and accuracy?}}
\item {\bf RQ2.} %\sout{How do users make use of the aggregate column results in conjunction with the raw spreadsheet data to analyze data?} 
\saj{How did \noah and it's components affect spreadsheet users'
navigational experiences?}
%\item {\bf RQ3.} \sout{How does bin customization help users inject domain knowledge in addressing their specific exploration questions compared to the automated overview?} \sout{\saj{How do users utilize additional features of \noah, \eg bin customization, during navigation?}
%\item {\bf RQ4.} \sout{How is the user's subjective satisfaction with \noah compared to traditional spreadsheets?}}  \sout{\saj{How is the user's subjective satisfaction with \noah compared to the traditional spreadsheets when performing navigational tasks?}}
\squishend

>>>>

In this section, we present the design of a user study
to evaluate whether 
\noah helps address spreadsheet navigational challenges.
\subsection{Study Design and Participants}
\label{sec:study_design}
\newsaj{The goal of our study is to evaluate whether \noah addresses the navigational challenges in traditional spreadsheets. Furthermore, we wanted understand how the presence of the features introduced by a plugin like \noah impacts a spreadsheet user's navigation experience as they explore and make sense of data. Therefore, we decided to compare a \noah-integrated spreadsheet with a typical, popular one, Excel, for various navigational tasks in the spreadsheet domain.} \saj{Similar domain specific-evaluations have been performed for measuring the performance of various  overview+detail interfaces, \eg database browsing~\cite{north2000snap} or 
tree navigation~\cite{beard1990navigational}. As explained in Section~\ref{sec:related}, the goals \newsaj{and user populations} of spreadsheets and TDA tools are quite different. Therefore, we did not consider TDA tools for the comparative study.}
Our study was designed to answer the following questions:
\agp{The RQs feel repetitive. What did Karrie think.}
\newsaj{She was fine with it. I think there are subtle differences. the RQ1 tries corresponds to quantifiable performance metric like accuracy, latency. RQ2 is qualitative feedback of user experience.}
\squishlist
\item {\bf RQ1.} %\sout{How does the addition of an overview help address spreadsheet navigational challenges, with respect to quantitative performance metrics such as speed and accuracy?}
\saj{How did the addition of an overview like \noah impact \newsaj{a user's spreadsheet navigation performance  measured in terms of navigation task submission speed and accuracy and quantitative survey metrics such as speed of use?}}
\item {\bf RQ2.} %\sout{How do users make use of the aggregate column results in conjunction with the raw spreadsheet data to analyze data?} 
\saj{How did \noah and it's components affect spreadsheet users'
navigational experiences?}
%\item {\bf RQ3.} \sout{How does bin customization help users inject domain knowledge in addressing their specific exploration questions compared to the automated overview?} \sout{\saj{How do users utilize additional features of \noah, \eg bin customization, during navigation?}
%\item {\bf RQ4.} \sout{How is the user's subjective satisfaction with \noah compared to traditional spreadsheets?}}  \sout{\saj{How is the user's subjective satisfaction with \noah compared to the traditional spreadsheets when performing navigational tasks?}}
\squishend
